---
title: "Data Analysis 01 Template"
author: Kelly Livingston, Allison Mollenkamp, Carmen Molina Acosta, Chris Barylick
date: "3/2/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Out of control
17 years. 179 victims. The deadly toll of street racing in Los Angeles (https://www.latimes.com/projects/la-me-street-racing/).

## Install 

```{r}

install.packages ("sf")

```

## Load libraries

Loading required libraries for this analysis.

```{r}
# Load the tidyverse
library(tidyverse)

library(sf)

```

## Load and Cleaning Data

[In this block of text, describe the source of the data, and write a basic data dictionary for data you are working with]

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

street_racing_deaths <- read.csv("data/output/fatalities.csv")

hoods <- st_read("data/input/hoods.shp") %>% as_tibble() %>%
  select(-geometry)

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Valentina D’Alessandro] was one of at least 179 people who have died in Los Angeles County since 2000 in accidents where street racing was suspected, according to a Los Angeles Times analysis of coroner’s records, police reports and media accounts from 2000 to 2017.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

total_deaths <- street_racing_deaths %>%
  summarise(total_deaths = n())

# Display results of code below this codeblock

total_deaths

```

### Sentence 2

* **Sentence text**: The deadliest year on record was 2007, with 18 fatal crashes. After a period with relatively few recorded deaths, the count grew in recent years, with 15 fatalities in 2015, 11 in 2016 and 12 deaths in 2017, the Times analysis shows.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

deaths_by_year 

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-